---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ShipApp

<!-- badges: start -->
<!-- badges: end -->

The goal of ShipApp is to visualize on the map the longest distance between two consecutive observations for every ship in given dataset. The app marks two points on the map. When clicking a marker user is able to see detailed info about the cruise (i.e. ship name, geocoordinates and sailed distance in meters). Additionally the app enables user to filter ships by type and name.

## Availability

The app is available for user on [shinyapps.io](https://pionow8.shinyapps.io/ShipApp/) website.

## Installation

You can install the released version of ShipApp version from [GitHub](https://github.com/) with:

``` r
## install.packages("devtools")
devtools::install_github("pionow8/Ship_app")
```
## Example

This is a basic example which shows you how to solve a common problem:

```{r example, eval = FALSE}
library(ShipApp)
## basic example code
ShipApp::run_app()
```

## Data preprocessing

```{r, eval = FALSE}
library("data.table")
library("rworldmap")
library("geosphere")

## read ship data
dt <- data.table(read.csv2("~/ships.csv", 
                           sep = ",", 
                           header = TRUE,
                           encoding = "UTF-8"))

## add coordinates of previous location
dt[, LAT0:= c(NA, LAT[-.N]), by=SHIP_ID]
dt[, LON0:= c(NA, LON[-.N]), by=SHIP_ID]

## change lon and lat varibles to numeric
dt$LAT <- as.numeric(dt$LAT)
dt$LON <- as.numeric(dt$LON)
dt$LAT0 <- as.numeric(dt$LAT0)
dt$LON0 <- as.numeric(dt$LON0)

## count distance with Haversine methode
dt$DISTANCE <- distHaversine(p1 = dt[, c('LON0', 'LAT0')],
                             p2 = dt[, c('LON', 'LAT')])

## choose max observation for each ship
dt <- unique(dt[dt[, .I[DISTANCE == max(DISTANCE, na.rm = TRUE)], by=SHIP_ID]$V1])

## change DATETIME variable to date
dt$DATETIME <- as.POSIXct(dt$DATETIME)

## sort by SHIP_ID and DATETIME variables
dt <- dt[order(SHIP_ID, -DATETIME)]

## take only the most recent observation for each ship
dt <- dt[, head(.SD, 1), by = "SHIP_ID"]

saveRDS(dt, "~/Ships_Final.RDS")
```

## Conclusions

The app should be a helpful tool for port employees. Enabling them to see ah-hoc ship localization and be a good tool for further analyses.
